\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-19 arch.tex
further updates to arch

\subsubsection{Synchronization and Identifying Internal State
  Evolution}
\label{sec:arch:synch}

@@

Deliberation inside a reactor is based on \eu capabilities described
in Section \ref{sec:europa:arch}. However deliberation within \rx
comes with some augmentation -- deliberation information one reactor
provide to the architecture such as its deliberation latency and its
plan look-ahead both expressed in ticks -- and the fact that we
potentially need to interleave deliberation with not just the
synchronization process but also with deliberation of other
reactors\footnote{The current implementation of \rx is running on a
  single process and it is the responsibility of the agent itself to
  emulate reactor multi-threading for deliberation and
  synchronization.}. For this reason while the overall planning of a
deliberative reactor strictly relies on the \eu framework planning
solver, it slices the execution of Algorithm \ref{alg:europa:solve} in
atomic steps. The interruption is done around the recursive call (line
\ref{li:recurse}) while ensuring that the current partial plan is
never proven inconsistant at any single step. This can be managed by
allowing the call to backtrack until a \textsf{DecisionPoInt} that is
not exhausted (\ie $decision.hasNext()$ is \texttt{true}) is found in
the call stack.

A critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not inconsistent. % Indeed, it is always
% possible that the next call made by the agent is for synchronization
% which could fail if the $plan$ is inconsistent before any attempt is
% made for synchronization.
If the partial plan is found to be inconsistent during a deliberation
step we either backtrack in the decision stack until a node of the
tree is found with an alternate solution or the plan is relaxed since
there is no other alternative solution. We do this by removing all
goals and keeping only recent observations. % These constraints are yet
% again related to the fact that for the agent the only crucial
% functionality of a reactor is to resolve successfully its
% synchronization.

{\em Synchronization} and the dispatching of new {\em goals} from
\rx can occur between two steps and alter the plan. These alterations
exogenous to the deliberation process are dynamically impacting the set
of flaws that need to be resolved. Obviously new flaws are created
when a new goal is received from \rx and conversely often the recall
of an objective will remove the corresponding flaws\footnote{although
  if the flaws of this goal were previously resolved, the removal of
  this goal may create new flaws}. The impact of {\em synchronization}
can be more subtle as its alteration of the plan at the execution
frontier is much more drastic. We can still give a non exhaustive list
of possible alterations of the flaws directly related to {\em
  synchronization}:
\begin{enumerate}
\item Previous flaws directly related to the execution frontier have
  been resolved by the {\em synchronization} solver.
\item The {\em synchronization} solver may have created new flaws that
  were outside of its scope but are now part of the {\em deliberation}
  solver scope.
\item Last observations proved the current plan to be invalid
  requiring the {\em synchronization} solver to relax the existing
  plan in order to identify its current {\em external} state. (This
  case is developed further on the next section)
\end{enumerate}
Finally as {\em synchronization} occurs when the {\em
  execution frontier} advances, it implies that the reactor planning
scope is further in the future. This need to be reflected by the {\em
  deliberation} solver which need a new deliberation scope one tick
further in the future.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window}
  \caption{\small The planning window of the deliberation solver and its
    relation with the {\em dispatch window}. The dark grey
    tokens   with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$. All the other
    tokens (in {\color{green} green}) -- along with the goals received from the
    agent (in {\color{blue} blue}) which can overlap this window are
    evaluated.}
  \label{fig:plan:window}
\end{figure}

As Fig. \ref{fig:plan:window} illustrates,  such a
scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable''
future. This filtering window complements the \rx {\em dispatch
  window} from definition \ref{def:dispatch} which already filters
out at the agent level the goals that are before the reactor execution
latency or too far in the future. The planning window inside of the
reactor refines further this filtering as tokens can be excluded
dynamically from deliberation as their insertion in the plan and
scheduling with other tokens generated can reduced their temporal
scope outside of the planning scope. This helps the deliberation 
solver reduce the number of tokens to be evaluated and therefore
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

>>>>

\subsubsection{Synchronization and Identifying Internal State}
\label{sec:arch:synch}

@@

Deliberation inside a reactor is based on \eu capabilities described
in Section \ref{sec:europa:arch}. However this deliberation comes with
some augmentation namely the reactor's deliberation latency and its
plan look-ahead both expressed in ticks and that the agent needs to
interleave deliberation with not just the synchronization but also
with deliberation of other reactors\footnote{The current
  implementation of \rx is running on a single process and it is the
  responsibility of the agent itself to emulate reactor
  multi-threading for deliberation and synchronization.}.  For this
reason while the reactor relies on the \eu solver, it slices the
execution of Algorithm \ref{alg:europa:solve} into atomic steps. This
interruption occurs around the recursive call line \ref{li:recurse}
while ensuring that the current partial plan is not inconsistant at
any single step. This is managed by allowing the call to backtrack
until a \textsf{DecisionPoInt} that is not exhausted (\ie
$decision.hasNext()$ is \texttt{true}) is found in the call stack.

A critical aspect of this deliberation step is to ensure that at the
exit of this call the $plan$ is not inconsistent. % Indeed, it is always
% possible that the next call made by the agent is for synchronization
% which could fail if the $plan$ is inconsistent before any attempt is
% made for synchronization.
If the partial plan is found to be inconsistent during a deliberation
step we either backtrack in the decision stack until a node of the
tree is found with an alternate solution or the plan is relaxed since
there is no other alternative solution. We do this by removing all
goals while keeping only recent observations.

{\em Synchronization} and the dispatching of new {\em goals} from \rx
can occur between two steps and alter the plan. These alterations
exogenous to the deliberation process dynamically impact the set of
flaws that need to be resolved. New flaws are created when a new goal
is received from \rx and conversely the recall of an objective could
remove the corresponding flaws\footnote{However if the flaws related
  to this goal were resolved, the removal of this goal may likely
  create new flaws.}. The impact of {\em synchronization} can be more
subtle as its alteration of the plan at the execution frontier is more
drastic. We can still give a non exhaustive list of possible
alterations of the flaws directly related to {\em synchronization}:

\item Previous flaws directly related to the execution frontier have
  been resolved by the {\em synchronization} solver.

\item The {\em synchronization} solver may have created new flaws that
  were outside of its scope but are now part of the scope of the {\em
    deliberation} solver.

\item The last observation has proved that the current plan is invalid
  requiring the {\em synchronization} solver to relax the existing
  plan to identify its current {\em external} state. (This case is
  developed further on the next section).

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window}
  \caption{\small The planning window of the deliberation solver and
    its relation to the {\em dispatch window}. The dark grey tokens
    with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$. All the other
    tokens ({\color{green} green}) along with the goals received from
    the agent ({\color{blue} blue}) which can overlap this window, are
    evaluated.}
  \label{fig:plan:window}
\end{figure}

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch
  window} from Definition \ref{def:dispatch} which filters out at the
agent level those goals that are before the reactor execution latency
or too far into the future. The planning window inside of a reactor
refines this filtering further, as tokens can be excluded dynamically
from deliberation as their insertion in the plan and scheduling with
other tokens generated, can reduce their temporal scope outside of the
planning scope \kcomment{ this last sentence should be reworked}. This
helps the deliberation solver reduce the number of tokens to be
evaluated reducing deliberation cost and increasing performance. As
time advances, flaws that were initially beyond this temporal horizon
will become active which a reactor can evaluate. This results in an
apparent \emph{continuous planning} of the reactor. When there are no
more flaws present in the current planning scope a partial plan
solution is found. This has two effects:
